---
title: "hw4"
author: "Daoyang E"
date: "10/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tibble)
tibble2 <- tribble(
  ~i, ~X_i,  ~Y_i,
  1, 1, 16,
  2, 0, 9,
  3, 2, 17,
  4, 0, 12,
  5, 3, 22,
  6, 1, 13,
  7, 0, 8,
  8, 1, 15,
  9, 2, 19,
  10, 0, 11
)
```

```{r}
X <- matrix(c(1,1,1,1,1,1,1,1,1,1,1,0,2,0,3,1,0,1,2,0), nrow = 10, ncol = 2)
Y <- matrix(c(16, 9,17,12,22,13,8,15,19,11), nrow = 10, ncol = 1)
```

```{r}
X_trans <- t(X)
```

### a 1
```{r}
library(matlib)
inv(X_trans %*% X)
```

Answer for a(1) is $\begin{pmatrix}0.2&-0.1\\ -0.1&0.1\end{pmatrix}$

### a 2

```{r}
b <- inv(X_trans %*% X) %*% X_trans %*% Y
b
```

Answer for a(2) is $\begin{pmatrix}10.2\\ 4.0\end{pmatrix}$

### a 3

```{r}
e <- Y - X %*% b
e
```

Answer for a(3) is $\begin{pmatrix}1.8\\ -1.2\\ -1.2\\ 1.8\\ -0.2\\ -1.2\\ -2.2\\ 0.8\\ 0.8\\ 0.8\end{pmatrix}$

### a 4

```{r}
H <- X %*% inv(X_trans %*% X) %*% X_trans
H
```

Answer for a 4 is just denoted in the R output. H is a 10 * 10 matrix.

### a 5

```{r}
SSE <- t(e) %*% e
SSE
```

The answer for a(5), SSE is 17.6

### a 6

```{r}
model1 <- lm(Y ~ X, tibble2)
anova(model1)
```

I see MSE in anova table is 2.2.

```{r}
MSE <- 2.2
s_square_b <- MSE * inv(X_trans %*% X)
s_square_b
```

The answer for a(6) is $\begin{pmatrix}0.44&-0.22\\ -0.22&0.22\end{pmatrix}$

### a 7

```{r}
X_h <- matrix(c(1,2), nrow = 2, ncol = 1)
X_h_tran <- t(X_h)
Y_h <- X_h_tran %*% b
Y_h
```

The answer for a(7), Y_h here is 18.2

### a 8

```{r}
s_square_yh <- MSE * (X_h_tran %*% inv(X_trans %*% X) %*% X_h)
s_square_yh
```

The answer for a(8) is 0.44

### b 1

$s^2\left\{b_1\right\}$ is equal to 0.22

### b 2

$s\left\{b_0,b_1\right\}$ is -0.22

### b 3

```{r}
sqrt(0.44)
```

$s\left\{b_0\right\}$ is 0.663


### c

```{r}
J <- matrix(rep(1, 100), nrow = 10, ncol = 10)
SSR <- H - 0.1*J
SSR
```

The matrix of Quadratic form for SSR is shown above