---
title: "hw8"
author: "Daoyang E"
date: "11/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(readr)
library(tidyr)
library(matlib)
library(modelr)
CH06PR15 <- read_csv("C:/Users/edaoy/Desktop/5205/CH06PR15.txt", 
    col_names = FALSE)
```


```{r}
ch6 <- CH06PR15 %>% separate(X1, into = c("Yi", "Xi1", "Xi2", "Xi3"), sep = "   ")
```


```{r}
ch6$Yi <- as.numeric(ch6$Yi)
ch6$Xi1 <- as.numeric(ch6$Xi1)
ch6$Xi2 <- as.numeric(ch6$Xi2)
ch6$Xi3 <- as.numeric(ch6$Xi3)
```


```{r}
X <- ch6[,2:4]
X <- as.matrix(X)
X_trans <- t(X)
H <- X %*% inv(X_trans%*%X) %*% X_trans
```


### a

```{r}
model <- lm(Yi~Xi1 + Xi2 + Xi3, data = ch6)
residuals <- summary(model)$residuals
ch6 <- cbind(ch6, residuals)
ch6$hii <- 0
for(i in 1:46){
  ch6$hii[i] = H[i,i]
}

anova(model)
SSE <- anova(model)[4,2]
```

```{r}
library(dplyr)
ch6 <- ch6 %>% mutate(ti = residuals * sqrt(41/(SSE*(1-hii)-residuals^2)))
```

```{r}
qt(1-0.1/(2*46),41)
```

```{r}
ch6 %>% filter(ti > 3.27)
```

The null hypothesis is that $H_0$:there is no outlier.
The decision rule is using bonferroni simultaneous test procedure with a= 0.1.
Since I see that there is no studentized deleted residuals greater than the test statistics 3.27.
So I concluded that there is no outlying Y observations

### b

```{r}
h_hat <- 4/46
ch6 %>% filter(hii >2 * h_hat)
```

I see that there is no X leverage value greater than 2p/n, so I conclude that there is no outlying X values.

### c

```{r}
X_new <- c(30,58,2)
t(X_new)%*%inv(t(X)%*%X)%*%X_new
```

```{r}
t(X_new)%*%inv(t(X)%*%X)%*%X_new > 2*h_hat
```

The new h is much greater than the leverage values, indicating a hidden extrapolation

### d

```{r}
#DFFITS

#case11
ch6$ti[11] * sqrt(ch6$hii[11]/(1-ch6$hii[11]))
#case17
ch6$ti[17] * sqrt(ch6$hii[17]/(1-ch6$hii[17]))
#case27
ch6$ti[27] * sqrt(ch6$hii[27]/(1-ch6$hii[27]))
```

For DFFITS criteria, none of case 11, 17, 27 are influential cases

```{r}
#cooks
MSE <- anova(model)[4,3]
#case11
(ch6$residuals[11]^2)/(4 * MSE)*(ch6$hii[11]/(1-ch6$hii[11])^2)
#case 17
(ch6$residuals[17]^2)/(4 * MSE)*(ch6$hii[17]/(1-ch6$hii[17])^2)
#case 27
(ch6$residuals[27]^2)/(4 * MSE)*(ch6$hii[27]/(1-ch6$hii[27])^2)
```
All the cook's distance value is very small, inficating no influential cases

```{r}
#DFBETAS

dfbetas <- as.data.frame(dfbetas(model))
dfbetas[c(11,17,27),]
```

There is no beta exceeding 1 in DFBETAS case, concluding that none of the three cases is influential

### e
```{r}
#without case 11
wo11 <- ch6[-11,]
model11 <- lm(Yi~Xi1 + Xi2+ Xi3, data = wo11)
coef11<- model11$coefficients
ch6 <- ch6 %>% mutate(Y11 = coef11[1]+coef11[2]*Xi1 + coef11[3]*Xi2+coef11[4]*Xi3) 
sum((ch6$Y11-ch6$Yi)/ch6$Yi)*100/46

```

```{r}
#without case 17
wo17 <- ch6[-17,]
model17 <- lm(Yi~Xi1 + Xi2+ Xi3, data = wo17)
coef17<- model17$coefficients
ch6 <- ch6 %>% mutate(Y17 = coef17[1]+coef17[2]*Xi1 + coef17[3]*Xi2+coef17[4]*Xi3) 
sum((ch6$Y17-ch6$Yi)/ch6$Yi)*100/46
```

```{r}
#without case 27
wo27 <- ch6[-27,]
model27 <- lm(Yi~Xi1 + Xi2+ Xi3, data = wo27)
coef27<- model27$coefficients
ch6 <- ch6 %>% mutate(Y27 = coef27[1]+coef27[2]*Xi1 + coef27[3]*Xi2+coef27[4]*Xi3) 
sum((ch6$Y27-ch6$Yi)/ch6$Yi)*100/46
```

None of the percentage is above 5 percent, meaning that they are not influential cases

### f
```{r}
ch6 <- ch6 %>% mutate(cook = (residuals^2)/(4 * MSE)*(hii/(1-hii)^2))
ch6$index = 1:46
```


```{r}
ggplot(data = ch6, aes(x = index, y = cook))+
  geom_point()+
  geom_line()+
  labs(title = "Index Influence Plot",
        x = "Cook's Distance D",
        y = "Case Index Number")
```

